---
title: "Lab 6 Comparing two means"
author: "Kelsey Leach"
date: "`r Sys.Date()`"
output: github_document
---

Researchers studying the number of electric fish species living in various parts of the Amazon basin were interested in whether the presence of tributaries affected the local number of electric fish species in the main rivers (Fernandes et al. 2004).

They counted the number of electric fish species above and below the entrance point of a major tributary at 12 different river locations.

The data is provided in your GitHub repository.

For each question below, write a sentence answering the question and show the code you used to come up with the answer, if applicable.

```{r Loading the data, include=FALSE}
library(tidyverse)
fish <- read_csv("chap12q19ElectricFish.csv")

fish_long <- 
  pivot_longer(fish, speciesUpstream:speciesDownstream,
               names_to = "location",
               values_to = "species") %>% 
  mutate(location = str_remove(location, c("species"))) 

```


## Question A

> What is the mean difference in the number of species between areas upstream and downstream of a tributary? What is the 95% confidence interval of this mean difference. Show your code and write a sentence giving your answer

The mean difference in the number of species between the areas is 1.83334 and the 95% confidence interval is -4.587031 to 8.253697.

```{r}

t.test(formula = species ~ location, data = fish_long)
```

## Question B

> Test the hypothesis that the tributaries have no effect on the number of species of electric fish.

We fail to reject the null hypothesis the the tributaries have no effect on the number of species of electric fish. This is because the difference between the two locations is not statistically significant (p > 0.05).

```{r}
t.test(formula = species ~ location, data = fish_long)
```

## Question C

> State the assumptions that you had to make to complete parts (A) and (B). Create a graph to assess whether one of those assumptions was met.

When performing a two-sample t-test, we assumed the measurements were taken from a random sample, the number of species is normally distributed, and the standard deviation of the number of species is the same in both samples. From the graph below, it does not appear that the normal distribution assumption was fulfilled. 

```{r}
fish_long %>% 
  ggplot(aes(x = species)) +
  geom_histogram(
    aes(fill = location), 
    bins = 15, 
    alpha = 0.5, 
    position = "identity"
  ) +
  #scale_fill_manual(values = c("darkorange","cyan4")) +
  theme_minimal()
```

## ANOVA

Fiddler crabs are so called because males have a greatly enlarged "major" claw, which is used to attract females and to defend a burrow.

Darnell and Munguia (2011) recently suggested that this appendage might also act as a heat sink, keeping males cooler while out of the burrow on hot days.

To test this, they placed four groups of crabs into separate plastic cups and supplied a source of radiant heat (60-watt light bulb) from above. The four groups were intact male crabs, male crabs with the major claw removed; male crabs with the other (minor) claw removed (control); and intact female fiddler crabs.

They measured the body temperature of crabs every 10 minutes for 1.5 hours. These measurements were used to calculate a rate of heat gain for every individual crab in degrees C/log minute. Rates of heat gain for all crabs are provided in the accompanying data file.

### Question D

Graph the distribution of body temperatures for each crab type:

```{r}
crab <- read_csv("chap15q27FiddlerCrabFans.csv")
crab <- na.omit(crab)
crab %>% 
  ggplot(aes(x = bodyTemperature)) +
  geom_histogram(
    aes(fill = crabType), 
    bins = 15, 
    alpha = 0.5, 
    position = "identity",
    na.rm = TRUE
  ) +
  scale_fill_manual(values = c("darkorange", "darkorchid", "cyan4", "deeppink")) +
  theme_minimal()

```

### Question E

Does body temperature varies among crab types? State the null and alternative hypothesis, conduct an ANOVA, and interpret the results.

The null hypothesis is that mean body temperature is equal between the crab types. The alternative hypothesis is that mean body temperature is different in at least one crab type. We reject the null hypothesis and accept the alternative hypothesis (p<0.05). There is a statistically significant difference between the mean body temperature in at least one of the crab types and the mean body temperature of the other types. 

```{r}
aov_crab_claw <-
  aov(bodyTemperature ~ crabType, data = crab)
aov_crab_claw
summary(aov_crab_claw)
```
