---
title: "Lab 6 Comparing two means"
author: "Kelsey Leach"
date: "`r Sys.Date()`"
output: github_document
---

Researchers studying the number of electric fish species living in various parts of the Amazon basin were interested in whether the presence of tributaries affected the local number of electric fish species in the main rivers (Fernandes et al. 2004).

They counted the number of electric fish species above and below the entrance point of a major tributary at 12 different river locations.

The data is provided in your GitHub repository.

For each question below, write a sentence answering the question and show the code you used to come up with the answer, if applicable.

```{r Loading the data, include=FALSE}
library(tidyverse)
fish <- read_csv("chap12q19ElectricFish.csv")

fish_long <- 
  pivot_longer(fish, speciesUpstream:speciesDownstream,
               names_to = "location",
               values_to = "species") %>% 
  mutate(location = str_remove(location, c("species"))) 

```


## Question A

> What is the mean difference in the number of species between areas upstream and downstream of a tributary? What is the 95% confidence interval of this mean difference. Show your code and write a sentence giving your answer

The mean difference in the number of species between the areas is 1.83334 and the 95% confidence interval is -4.587031 to 8.253697.

```{r}

t.test(formula = species ~ location, data = fish_long)
```

## Question B

> Test the hypothesis that the tributaries have no effect on the number of species of electric fish.

We fail to reject the null hypothesis the the tributaries have no effect on the number of species of electric fish. This is because the difference between the two locations is not statistically significant (p > 0.05).

```{r}
t.test(formula = species ~ location, data = fish_long)
```

## Question C

> State the assumptions that you had to make to complete parts (A) and (B). Create a graph to assess whether one of those assumptions was met.

When performing a two-sample t-test, we assumed the measurements were taken from a random sample, the number of species is normally distributed, and the standard deviation of the number of species is the same in both samples. From the graph below, it does not appear that the normal distribution assumption was fulfilled. 

```{r}
fish_long %>% 
  ggplot(aes(x = species)) +
  geom_histogram(
    aes(fill = location), 
    bins = 15, 
    alpha = 0.5, 
    position = "identity"
  ) +
  scale_fill_manual(values = c("darkorange","cyan4")) +
  theme_minimal()
```
